---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE)
```

# scrutr

```{=html}
<!-- badges: start -->
```

[![CRAN status](https://www.r-pkg.org/badges/version/scrutr)](https://CRAN.R-project.org/package=scrutr)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/scrutr)](https://cran.r-project.org/package=scrutr)
[![R-CMD-check](https://github.com/danielrak/scrutr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/danielrak/scrutr/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/danielrak/scrutr/branch/main/graph/badge.svg)](https://app.codecov.io/gh/danielrak/scrutr?branch=main)
```{=html}
<!-- badges: end -->
```

> `scrutr` was previously developed under the name `industtry` (2024-2025) and has been renamed ahead of its first CRAN release.

`scrutr` is an R toolkit for **scrutinizing collections of structured datasets** (data frames). It targets workflows where you need to apply the same inspection, profiling, or conversion procedure across many related tables.

Two guiding ideas:

1. **"Collections first"**: common tasks (inspecting, detecting schema differences, batch conversion) operate at the *collection* level, not one dataset at a time.
2. **"Operational tooling"**: lightweight helpers for paths, duplicates, join checks, and string replacements that recur in production data pipelines.

---

## Installation

```r
# install.packages("devtools")
devtools::install_github("danielrak/scrutr")
```

---

## API overview

```{r, echo = FALSE}
library(scrutr)
api_map <- tibble::tribble(
  ~Area, ~`Main intent`, ~`Key functions`,
  "Inspection & profiling",
  "Profile one dataset or a whole folder; export diagnostics to Excel.",
  "inspect(), inspect_write(), inspect_vars()",
  "Schema detection",
  "Detect variable presence across datasets; compare classes.",
  "vars_detect(), vars_compclasses(), detect_chars_structure_datasets()",
  "Batch conversion / renaming",
  "Convert file formats or rename files at scale via Excel masks.",
  "convert_r(), convert_all(), mask_convert_r(), mask_rename_r(), rename_r()",
  "Data hygiene",
  "Duplicate diagnostics, join checks, proportions.",
  "dupl_show(), dupl_sources(), ljoin_checks(), table_prop()",
  "Paths & filesystem",
  "Replicate folder structures, move through paths.",
  "folder_structure_replicate(), path_move()",
  "Utilities",
  "Batch string replacement.",
  "replace_multiple()"
)
knitr::kable(api_map)
```

---

## Quick examples

### Inspect a single data frame

```{r}
library(scrutr)
result <- inspect(CO2)
head(result)
```

### Inspect a folder of datasets and export to Excel

```{r, eval = FALSE}
mydir <- file.path(tempdir(), "example")
dir.create(mydir, showWarnings = FALSE)

saveRDS(cars, file.path(mydir, "cars.rds"))
saveRDS(mtcars, file.path(mydir, "mtcars.rds"))

inspect_vars(
  input_path = mydir,
  output_path = mydir,
  output_label = "diagnostics",
  considered_extensions = "rds"
)
```

### Batch conversion with `convert_all()`

```{r, eval = FALSE}
convert_all(
  input_folderpath = mydir,
  considered_extensions = "rds",
  to = "csv",
  output_folderpath = file.path(mydir, "csv_output")
)
```

### Mask-driven conversion with `convert_r()`

```{r, eval = FALSE}
# 1. Generate an Excel mask template
mask_convert_r(output_path = mydir)

# 2. Fill in the mask, then run:
convert_r(
  mask_filepath = file.path(mydir, "mask_convert_r.xlsx"),
  output_path = mydir
)
```
